The 18.821 Mathematics Project Lab Report [Proofs]

### 2013-09-27 paper.tex
Finishing up the symmetrical quadrilateral proof.

\begin{lemma}
  If $f(t), g(t) > 0$ for all $t$, then $\argmin_t \sqrt{f(t)} + \sqrt{g(t)} = \argmin_t f(t) + g(t)$.
  \label{lemma:minsqrt}
\end{lemma}
\proof Let $t_1, t_2$ be such that $f(t_1) + g(t_1) < f(t_2) + g(t_2)$. Then we can show that $\sqrt{f(t_1)} + \sqrt{g(t_1)} < \sqrt{f(t_2)} + \sqrt{g(t_2)}$. To do this, let's define $k(t) = t^{1/2}$ as the square root function. Then $\sqrt{f(t))} = k(f(t))$. Now we know that $k(t)$ is a strictly monotonically increasing function when $t > 0$, which means that $k(x) < k(y)$ if and only if $x < y$ (assuming we can confine $x, y$ to be non-negative).

Because this is the case, we see that $k(f(x)) < k(f(y))$ if and only if $f(x) < f(y)$ (the same goes for $g$). Thus, we see that if we have found the minimum $t_m$ to $\min_t k(f(t)) + k(g(t))$, then it is the case that $k(f(t_m)) + k(g(t_m)) < k(f(t)) + k(g(t))$ for all $t \neq t_m$ (again where $t > 0$). Following our train of logic, we see that $f(t_m) + g(t_m) < f(t) + g(t)$ for all $t \neq t_m$, which means that $t_m$ is a minimum of $f(t) + g(t)$. Thus by finding a minimum $t_m$ to $k(f(t)) + k(g(t)) = \sqrt{f(t)} + \sqrt{g(t)}$, we also found a minimum to $f(t) + g(t)$. \qed

Since we've proven this lemma, we can invoke it upon $\min l_3 + l_4$.

>>>>

\begin{lemma}
  If $f(t), g(t) > 0$ and $k(t) > 0$ is a strictly monotonically increasing function for all $t$, then $\argmin_t k(f(t)) + k(g(t)) = \argmin_t f(t) + g(t)$.
  \label{lemma:minsqrt}
\end{lemma}
\proof Let $t_1, t_2$ be such that $f(t_1) + g(t_1) < f(t_2) + g(t_2)$. In this proof, we will show that $k(f(t_1)) + k(g(t_1)) < \sqrt{f(t_2)} + \sqrt{g(t_2)}$. Since $k(t)$ is a strictly monotonically increasing function when $t > 0$, we know that $k(x) < k(y)$ if and only if $x < y$ (assuming we can confine $x, y$ to be non-negative).

Because this is the case, we see that $k(f(x)) < k(f(y))$ if and only if $f(x) < f(y)$ (the same goes for $g$). Thus, we see that if we have found the minimum $t_m$ to $\argmin_t k(f(t)) + k(g(t))$, then it is the case that $k(f(t_m)) + k(g(t_m)) < k(f(t)) + k(g(t))$ for all $t \neq t_m$ (again where $t > 0$). Following our train of logic, we see that $f(t_m) + g(t_m) < f(t) + g(t)$ for all $t \neq t_m$, which means that $t_m$ is a minimum of $f(t) + g(t)$. Thus by finding a minimum $t_m$ to $k(f(t)) + k(g(t))$, we also found a minimum to $f(t) + g(t)$. \qed

Since we've proven this lemma, we can invoke it upon $\min l_3 + l_4$. Since $l_3 = \sqrt{z_3}$ and $l_4 = \sqrt{z_4}$, we can use $k(t) = \sqrt{t}$ and we can write $\min l_3 + l_4 = \min z_3 + z_4$ by using our lemma. Thus, we now want to solve the problem:
\begin{eqnarray}
  \argmin_{\theta} 2 \left(\frac{l_1}{2} - \frac{l_2}{2} \sin \theta \right)^2 + \left( d - \frac{l_2}{2} \cos \theta \right)^2 + \left( d + \frac{l_2}{2} \cos \theta \right)^2
\end{eqnarray}

Now, we can expand out our expression and use the fact that $\sin^2 \theta + \cos^2 \theta = 1$ to obtain a much simpler (but equivalent) minimization problem:
\begin{eqnarray}
  \argmin_{\theta} \frac{l_1^2}{2} + \frac{l_2^2}{2} + 2d^2 - l_1 l_2 \sin \theta
\end{eqnarray}

We note that $l_1, l_2,$ and $d$ are all constants which are given to us in the problem. Therefore, the minimization problem really boils down to
\begin{eqnarray}
  \argmin_{\theta} - l_1 l_2 \sin \theta = \argmax_{\theta} \sin \theta
\end{eqnarray}

Thus, we see that $\theta = \frac{\pi}{2}$ minimizes the perimeter of the quadrilateral in the symmetric case.
